---
title: "Tidy Tuesday"
subtitle: "Week 33"
author: "Cristian T"
date: last-modified
format: 
   html:
     df-print: paged
     embed-resources: true
editor: 
  markdown: 
    wrap: 72
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(warning = FALSE, message = FALSE)
```


```{r}
# Load the tidytuesday package
suppressMessages(library(tidytuesdayR))
suppressMessages(library(skimr))
suppressMessages(library(tidyverse))
suppressMessages(library(dplyr))


# Load the current week's dataset
tuesdata <- tidytuesdayR::tt_load('2024-08-13')

worlds_fairs <- tuesdata$worlds_fairs

# Explore the structure of the dataset
str(worlds_fairs)
skim(worlds_fairs)


# View the first few rows of the dataset
head(worlds_fairs)


#tidytuesdayR::use_tidytemplate()
```

#### Missing Values
Notables: 11 missing values (about 15.7% of the rows).
Visitors: 14 missing values (about 20% of the rows).
Cost: 34 missing values (about 48.6% of the rows).
Area: 5 missing values (about 7.1% of the rows).
Attending Countries: 5 missing values (about 7.1% of the rows).
Most of the cost fetched from https://jdpecon.com/expo/expolist.html

# Clean the data
```{r}

#selected_data <- worlds_fairs %>% dplyr::select("country", "start_year", "city", "cost")
#selected_data

# add missing values
# https://jdpecon.com/expo/expolist.html
worlds_fairs[70, "cost"] <- 14600 # Dubai 2022
worlds_fairs[69, "cost"] <- 3000 # Kazakhstan 2017
worlds_fairs[68, "cost"] <- 1640 # Italy 2015
worlds_fairs[65, "cost"] <- 1070 # Spain 2008
worlds_fairs[64, "cost"] <- 3130 # Japan 2005
worlds_fairs[62, "cost"] <- 2952 # Portugal 1998 | Estimated $1753 to $2398 million - We will use median 2952 rounded
worlds_fairs[61, "cost"] <- 2800 # South Korea 1993
worlds_fairs[60, "cost"] <- 568.75 # Italy 1992 $487.5 to $650 million according to the New York Times - We will use median 568.75 rounded
worlds_fairs[59, "cost"] <- 2332 # Spain 1992
worlds_fairs[54, "cost"] <- 763.7 # Japan 1985
worlds_fairs[50, "cost"] <- 304 # Japan 1975
worlds_fairs[12, "cost"] <- 1.14 # Belgium 1897


# Corrections
# https://jdpecon.com/expo/expolist.html
worlds_fairs[1, "cost"] <- 1.67 # London 1851 
worlds_fairs[2, "cost"] <- 5.7 # France 1855 
worlds_fairs[3, "cost"] <- 2.3 # London 1862 
worlds_fairs[4, "cost"] <- 4.5 # Paris 1867
worlds_fairs[5, "cost"] <- 12 # Vienna 1873
worlds_fairs[6, "cost"] <- 9 # Philadelphia 1876 
# worlds_fairs[7, "cost"] <- 11 # France 1878	
worlds_fairs[8, "cost"] <- 1.56 # Melbourne 1880
worlds_fairs[9, "cost"] <- 1.7 # Spain 1881
worlds_fairs[10, "cost"] <- 8.01 # France 1889 41.5 million French francs would be approximately 8.01 million USD
worlds_fairs[11, "cost"] <- 27.3 # Chicago 1893
worlds_fairs[12, "cost"] <- 1.14 # Belgium	1897
worlds_fairs[13, "cost"] <- 18.75 # Paris 1900 
worlds_fairs[14, "cost"] <- 26.5 # St Louis 1904
worlds_fairs[15, "cost"] <- 2.89 # Belgium 1905
worlds_fairs[16, "cost"] <- 2.6 # Milan 1906
worlds_fairs[17, "cost"] <- 3.55 # Belgium	1910
# worlds_fairs[18, "cost"] <- NA # Italy	1911
worlds_fairs[19, "cost"] <- 3.3 # Belgium	1913
worlds_fairs[20, "cost"] <- 24.7 # San Francisco	1915
# worlds_fairs[21, "cost"] <- 25 # Spain	1929
worlds_fairs[22, "cost"] <- 31 # Chicago 1933
worlds_fairs[23, "cost"] <- 22.4 # Belgium 1935 
# worlds_fairs[24, "cost"] <- NA # Sweden	1936
worlds_fairs[25, "cost"] <- 57.25 # Paris 1937 
# worlds_fairs[26, "visitors"] <- NA # Finland	1938
worlds_fairs[27, "cost"] <- 160 # New York City 1939
worlds_fairs[27, "attending_countries"] <- 52 # New York City 1939

# https://en.worldfairs.info/expodonnees.php?expo_id=32
worlds_fairs[28, "area"] <- 80 # Belgium	1939
worlds_fairs[28, "visitors"] <- 1.7 # Belgium	1939
worlds_fairs[28, "cost"] <- 58.51 # Belgium	1939 | 150 million Belgian francs would be approximately $58.51 million 2015 dollars 

# worlds_fairs[29, "cost"] <- NA # France	1947
# worlds_fairs[30, "cost"] <- NA # Sweden	1949
# worlds_fairs[31, "cost"] <- NA # France	1949

# https://www.bie-paris.org/site/en/latest/blog/entry/cultivating-tourism-in-haiti-the-bicentennial-international-exposition-of-port-au-prince-1949
worlds_fairs[32, "cost"] <- 4 # Expo Port-au-Prince 1949
worlds_fairs[32, "visitors"] <- 0.07 # Expo Port-au-Prince 1949
# worlds_fairs[33, "cost"] <- NA # France	1951
# worlds_fairs[34, "cost"] <- NA # Italy	1953
worlds_fairs[35, "visitors"] <- 0.6 # Expo 1953 Jerusalem
# worlds_fairs[36, "cost"] <- NA # Italy	1954
worlds_fairs[37, "visitors"] <- 0.12 # Italy	1955
# worlds_fairs[38, "cost"] <- NA # Sweden	1955
# worlds_fairs[39, "cost"] <- NA # Israel	1956
# worlds_fairs[40, "cost"] <- NA # West Germany	1957
worlds_fairs[41, "cost"] <- 43.4 # Belgium	1958
# worlds_fairs[42, "cost"] <- NA # Italy	1961
worlds_fairs[43, "cost"] <- 22.8 # Seattle 1962
# worlds_fairs[44, "cost"] <- NA # West Germany	1965
worlds_fairs[45, "cost"] <- 384.7 # 1967 Canada 415.920 CAD million would be approximately 384.7 million USD
# worlds_fairs[46, "cost"] <- 156.00 # San Antonio 1968
worlds_fairs[47, "cost"] <- 247.54 # Japan	1970
# worlds_fairs[48, "cost"] <- NA # Hungary	1971
worlds_fairs[49, "cost"] <- 78.40 # Spokane	1974
worlds_fairs[50, "cost"] <- 155.2 # Japan	1975
# worlds_fairs[51, "cost"] <- NA # Bulgaria	1981
worlds_fairs[52, "cost"] <- 111 # Knoxville	1982
worlds_fairs[53, "cost"] <- 442.5 # New Orleans	1984
worlds_fairs[54, "cost"] <- 763.7 # Japan	1985
# worlds_fairs[55, "cost"] <- NA # Bulgaria	1985
worlds_fairs[56, "cost"] <- 609.52 # Canada	1986
worlds_fairs[57, "cost"] <- 642.2 # Australia	1988
# worlds_fairs[58, "cost"] <- NA # Bulgaria	1991
# worlds_fairs[59, "cost"] <- 2332 # Spain 1992
# worlds_fairs[60, "cost"] <- 568.75 # Italy 1992 
# worlds_fairs[61, "cost"] <- 2800 # South Korea 1993
# worlds_fairs[62, "cost"] <- 2952 # Portugal 1998 
worlds_fairs[63, "cost"] <- 2238 # Germany	2000
# worlds_fairs[64, "cost"] <- 3130 # Japan 2005
# worlds_fairs[65, "cost"] <- 1070 # Spain 2008
worlds_fairs[66, "cost"] <- 8800  # China	2010

#https://www.exhibitionworld.co.uk/crossing-oceans-expo-2012-yeosu-korea
worlds_fairs[67, "cost"] <- 2000 # South Korea 2012

worlds_fairs[68, "cost"] <- 1640 # Italy 2015
worlds_fairs[69, "cost"] <- 3000 # Kazakhstan 2017
worlds_fairs[70, "cost"] <- 14600 # Dubai 2022



# we still have 21 null values for cost
# visitors	10
# area	4 and attending_countries	5	
```

To address the missing values, we will use an imputation method to help fill in the remaining missing values using a more sophisticated technique than the median. Using the mice package (https://amices.org/mice/), we will consider the relationships between variables to fill in the missing data.

Instead of using KNN, we decided to try PMM (Predictive Mean Matching). This matches each missing value with the nearest observed values and imputes from those values. Effective for maintaining the distribution of the data. PMM does multiple imputations by default when you set m > 1, creating multiple versions of the dataset with imputed values, reflecting the uncertainty of the missing data.
```{r}

suppressMessages(library(mice))

#  Show the missing data pattern
md.pattern(worlds_fairs) # Blue is observed, red is missing

imputed_data <- mice(worlds_fairs, method = 'pmm', m = 5, maxit = 50)
#imputed_data <- mice(worlds_fairs, method = 'norm', m = 5, maxit = 50) # Gives negative values
#imputed_data <- mice(worlds_fairs, method = 'norm.predict', m = 5, maxit = 50) # Gives negative values
#imputed_data <- mice(worlds_fairs, method = 'mean', m = 5, maxit = 50) # Gives negative values
#imputed_data <- mice(worlds_fairs, method = 'pmm', m = 5, maxit = 50, seed = 1842)
# we are using the Predictive Mean Matching method and we generated 5 different versions of the dataset, each with slightly different imputed values and 50 iterations with random seed 1842

# Inspect the imputed data.
#summary(imputed_data)

# Matrix of the imputed values for the cost variable
imputed_data$imp$cost

# Visual summary of the imputation process
stripplot(imputed_data)
# helps you understand how the imputed values compare to the observed data and whether the imputation seems reasonable.

# Extract the Dataset
worlds_fairs_imputed <- complete(imputed_data,3)

# Compare Before and After
summary(worlds_fairs$cost)
summary(worlds_fairs_imputed$cost)

#skim(worlds_fairs)
#skim(worlds_fairs_imputed)

imputed_fairs <- worlds_fairs_imputed %>% dplyr::select("start_year","country", "city", "cost")
imputed_fairs

```


```{r}
library(ggplot2)

ggplot(worlds_fairs_imputed, aes(x = start_year, y = visitors)) +
  geom_line() +
  labs(title = "Visitor Trends Over Time",
       x = "Year",
       y = "Visitors (in millions)") +
  theme_minimal()


```


```{r}
# Rename counties to 
worlds_fairs_imputed <- worlds_fairs_imputed %>%
  mutate(country = recode(country, "Colony of Victoria" = "Australia", "West Germany" = "Germany", "People's Republic of China" = "China",  "Austria-Hungary" = "Austria"))


ggplot(worlds_fairs_imputed %>%
         count(country) %>%
         arrange(n), 
       aes(x = reorder(country, n), y = n, fill = country)) +
  geom_bar(stat = "identity") +
  coord_flip() +
  labs(title = "Number of Expositions by Country",
       x = "Country",
       y = "Number of Expositions") +
  theme_minimal() +
  theme(legend.position = "none")

```

```{r}

cost_expositions_df <- worlds_fairs_imputed[, c("cost", "start_year")]

cost_expositions_df


worlds_fairs_imputed %>%
  group_by(start_year) %>%
  summarise(avg_cost = mean(cost, na.rm = TRUE)) %>%
  ggplot(aes(x = start_year, y = avg_cost)) +
  geom_line() +
  geom_point() +
  labs(title = "Average Cost of World Expositions Over Time",
       x = "Year",
       y = "Average Cost (in millions)") +
  theme_minimal()

```








```{r}
#sessioninfo::session_info(include_base = TRUE)
```

```{r}
ggplot(worlds_fairs_imputed, aes(x = cost, y = visitors)) +
  geom_point() +
  geom_smooth(method = "lm") +
  labs(title = "Cost vs. Visitors",
       x = "Cost (in millions)",
       y = "Visitors (in millions)") +
  theme_minimal()

cost_visitors_df <- worlds_fairs_imputed[, c("cost", "visitors")]

cost_visitors_df

```